---
title: "churnDataPreprocessing"
output: html_document
date: "2024-07-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dtplyr)
library(ggplot2)
library(tidyverse)
library(timeDate)

```

```{r}
mlData <- read.csv("cleanedDatasetV1.csv", header = T, stringsAsFactors = T)
```


```{r}
dim(mlData)
```
## Remove all rows with empty cells
```{r}
mlData <- na.omit(mlData)

```
`


## Deduce the Purchase/Not Purchase column from the target column

```{r}
mlData1 <- mlData%>%
mutate(Purchase = case_when(target == "p" ~  "Purchase", 
                              target == "n" ~  "No Purchase"))
```



## Change the date column from factor to date datatype
## mlData1$date <- as.Date(mlData1$date)


```{r}
mlData1$date <- as.Date(mlData1$date, format= "%d/%m/%Y")
```


```{r}
dim(mlData1)
```
## We have produced the Purchase column

```{r}


mlData1$Purchase <- as.factor(mlData1$Purchase)


```


## Create and add the Product category column
```{r}
mlData11 <- mlData1%>%
mutate(productCategory = case_when(asin == "1524763136" ~  "Book", 
                              asin == "B07PGL2N7J" ~  "Electronics",
                              asin == "B005OCFGTO" ~  "Movie",
                              asin == "B06X9T1Y3F" ~  "Curcumin",
                              asin == "B07XC6WRJQ" ~  "Food",
                              asin == "B088C2WD5F" ~  "Mask",
                              TRUE ~  "Perfumes"))
```


```{r}
dim(mlData11)
```


```{r}
mlData11$productCategory <- as.factor(mlData11$productCategory)
```




```{r}
mlData111 <- mlData11%>% group_by(productCategory)%>%mutate(categoryPurchase=sum(Purchase == "Purchase"))
```


```{r}
mlData1111 <- mlData111%>% group_by(productCategory)%>%mutate(categoryNoPurchase=sum(Purchase == "No Purchase"))
```


```{r}
mlData11111 <- mlData1111%>%
mutate(noPurchaseProbability = case_when(productCategory == "Book" ~ categoryNoPurchase/4999, 
                          productCategory == "Movie" ~ categoryNoPurchase/4999, 
                          productCategory == "Electronics" ~ categoryNoPurchase/4999,
                          productCategory == "Curcumin" ~ categoryNoPurchase/4999,
                          productCategory == "Food" ~ categoryNoPurchase/4999, 
                          productCategory == "Mask" ~ categoryNoPurchase/4187, 
                          productCategory == "Perfumes" ~ categoryNoPurchase/5000,
                         ))
```


```{r}
mlData11111$noPurchaseProbability <- round(mlData11111$noPurchaseProbability,2)
```




```{r}
write.csv(mlData11111,"modelData.csv")
```

